---
title: "Linear Transformations for the Independent Variable"
output: 
  html_document: 
    toc: true
---

```{r, message=FALSE, warning=FALSE}
library(gridExtra)
library(ggplot2)
library(dplyr)
```

## Information

The following plots were created on May 25, 2017 to understand linear transformations of the independent variable to to better fit linear relationships to non-linear data.

## Positive Values

### Linear Plot

```{r}
set.seed(123)
lin_data <- data.frame(x = runif(1000, 0, 100))
lin_data$y <- rnorm(1000, 10, 10) + 1.2 * lin_data$x
```

```{r}
mid <- ggplot(lin_data, aes(x,y)) + 
    geom_point() + 
    geom_smooth(method="loess", span = 0.3,color = "red") + 
    geom_smooth(method="lm", color = "blue") +
    theme(plot.title = element_text(hjust = 0.5)) +
    labs(title = "y = x")
```

### Linear-Reciprical

```{r}
# Left
left_data <- data.frame(x = lin_data$x)
left_data$y <- left_data$x # Change here
left_data$y <- left_data$y + rnorm(1000, 0.25*max(left_data$y), 0.125*max(left_data$y))

line <- ggplot(left_data, aes(x,y)) + 
    geom_point() + 
    geom_smooth(method="loess", span = 0.3,color = "red") + 
    geom_smooth(method="lm", color = "blue") +
    theme(plot.title = element_text(hjust = 0.8)) +
    labs(title = "y = x")

# Right
right_data <- data.frame(x = lin_data$x)
right_data$y <- 1/right_data$x # Change here
right_data$y <- right_data$y + rnorm(1000, 0.25*max(right_data$y), 0.125*max(right_data$y))
recip <- ggplot(right_data, aes(x,y)) + 
    geom_point() + 
    geom_smooth(method="loess", span = 0.3,color = "red") + 
    geom_smooth(method="lm", color = "blue") +
    labs(title = "y = 1/(x) \n <---1/(x) Transform----")
```

```{r, fig.width = 10}
grid.arrange(line,mid,recip, ncol = 3)
```

### Logistic-Exponential Regression

```{r fig.width = 9}
# Left 
left_data <- data.frame(x = lin_data$x)
left_data$y <- log(left_data$x) 
left_data$y <- left_data$y + rnorm(1000, 0.25*max(left_data$y), 0.125*max(left_data$y))

log <- ggplot(left_data, aes(x,y)) + 
    geom_point() + 
    geom_smooth(method="loess", span = 0.3,color = "red") + 
    geom_smooth(method="lm", color = "blue") +
    theme(plot.title = element_text(hjust = 0.8)) +
    labs(title = "y = log(x) \n ---Log(x) Transform--->")

# Right 
right_data <- data.frame(x = lin_data$x)
right_data$y <- exp(right_data$x)
right_data$y <- right_data$y + rnorm(1000, 0.25*max(right_data$y), 0.125*max(right_data$y))
exp <- ggplot(right_data, aes(x,y)) + 
    geom_point() + 
    geom_smooth(method="loess", span = 0.3,color = "red") + 
    geom_smooth(method="lm", color = "blue") +
    labs(title = "y = e^x \n <---Exp(x) Transform----")
```

```{r, fig.width = 10}
grid.arrange(log,mid,exp, ncol = 3)
```

### Square-Square Root

```{r}
# Left 
left_data <- data.frame(x = lin_data$x)
left_data$y <- left_data$x^2 # Change here
left_data$y <- left_data$y + rnorm(1000, 0.25*max(left_data$y), 0.125*max(left_data$y))

square <- ggplot(left_data, aes(x,y)) + 
    geom_point() + 
    geom_smooth(method="loess", span = 0.3,color = "red") + 
    geom_smooth(method="lm", color = "blue") +
    theme(plot.title = element_text(hjust = 0.8)) +
    labs(title = "y = x^2 \n ---x^2 Transform--->")

# Right 
right_data <- data.frame(x = lin_data$x)
right_data$y <- sqrt(right_data$x) # Change here
right_data$y <- right_data$y + rnorm(1000, 0.25*max(right_data$y), 0.125*max(right_data$y))
sqrt <- ggplot(right_data, aes(x,y)) + 
    geom_point() + 
    geom_smooth(method="loess", span = 0.3,color = "red") + 
    geom_smooth(method="lm", color = "blue") +
    labs(title = "y = sqrt(x) \n <---sqrt(x) Transform----")
```

```{r, fig.width = 10}
grid.arrange(square,mid,sqrt, ncol = 3)
```

### Cube-Cubic Root

```{r}
# Left 
left_data <- data.frame(x = lin_data$x)
left_data$y <- left_data$x^3 # Change here
left_data$y <- left_data$y + rnorm(1000, 0.25*max(left_data$y), 0.125*max(left_data$y))

cube <- ggplot(left_data, aes(x,y)) + 
    geom_point() + 
    geom_smooth(method="loess", span = 0.3,color = "red") + 
    geom_smooth(method="lm", color = "blue") +
    theme(plot.title = element_text(hjust = 0.8)) +
    labs(title = "y = x^3 \n ---x^3 Transform--->")

# Right 
right_data <- data.frame(x = lin_data$x)
right_data$y <- right_data$x^(1/3) # Change here
right_data$y <- right_data$y + rnorm(1000, 0.25*max(right_data$y), 0.125*max(right_data$y))
cubic_root <- ggplot(right_data, aes(x,y)) + 
    geom_point() + 
    geom_smooth(method="loess", span = 0.3,color = "red") + 
    geom_smooth(method="lm", color = "blue") +
    labs(title = "y = x^(1/3) \n <---x^(1/3) Transform----")
```

```{r, fig.width = 10}
grid.arrange(cube,mid,cubic_root, ncol = 3)
```



### Sin-Cos

```{r}
# Left 
left_data <- data.frame(x = lin_data$x)
left_data$y <- sin(left_data$x) # Change here
left_data$y <- left_data$y + rnorm(1000, 0.25*max(left_data$y), 0.125*max(left_data$y))

sin <- ggplot(left_data, aes(x,y)) + 
    geom_point() + 
    geom_smooth(method="loess", span = 0.3, color = "red") + 
    geom_smooth(method="lm", color = "blue") +
    theme(plot.title = element_text(hjust = 0.8)) +
    labs(title = "y = sin(x) \n ---sin(x) Transform--->")

# Right 
right_data <- data.frame(x = lin_data$x)
right_data$y <- cos(right_data$x) # Change here
right_data$y <- right_data$y + rnorm(1000, 0.25*max(right_data$y), 0.125*max(right_data$y))
cos <- ggplot(right_data, aes(x,y)) + 
    geom_point() + 
    geom_smooth(method="loess", span = 0.3, color = "red") + 
    geom_smooth(method="lm", color = "blue") +
    labs(title = "y = cos(x) \n <---cos(x) Transform----")
```

```{r test, fig.width = 10}
grid.arrange(sin,mid,cos, ncol = 3)
```

### Comparison Plot
```{r fig.height=15, fig.width=9}
one <- arrangeGrob(line,mid,recip, ncol = 3)
two <- arrangeGrob(log,mid,exp, ncol = 3)
three <- arrangeGrob(square,mid,sqrt, ncol = 3)
four <- arrangeGrob(cube,mid,cubic_root, ncol = 3)
five <- arrangeGrob(sin,mid,cos, ncol = 3)
grid.arrange(one, two, three, four, five,ncol = 1)
```

## Negative Values

### Valid Transforms:

* 1/x
* exp(x)
* x^2 - symmetric chape to positive
* x^3 - inverted shape to positive
* sin/cos 
